\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage[numbers]{natbib}

\usepackage{relsize}
\DeclareRobustCommand{\isactrlbsub}{\emph\bgroup\math{}\sb\bgroup\mbox\bgroup\isaspacing\itshape\smaller}
\DeclareRobustCommand{\isactrlesub}{\egroup\egroup\endmath\egroup}
\DeclareRobustCommand{\isactrlbsup}{\emph\bgroup\math{}\sp\bgroup\mbox\bgroup\isaspacing\itshape\smaller}
\DeclareRobustCommand{\isactrlesup}{\egroup\egroup\endmath\egroup}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

\input{header-delta-system}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{tt}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}

\renewcommand{\isacharunderscorekeyword}{\mbox{\_}}
\renewcommand{\isacharunderscore}{\mbox{\_}}
\renewcommand{\isasymtturnstile}{\isamath{\Vdash}}
\renewcommand{\isacharminus}{-}

\begin{document}

\title{Cofinality and the Delta System Lemma}
\author{
  Pedro S\'anchez Terraf\thanks{Universidad Nacional de C\'ordoba. 
    Facultad de Matem\'atica, Astronom\'{\i}a,  F\'{\i}sica y
    Computaci\'on.}
  \thanks{%
    Centro de Investigaci\'on y Estudios de Matem\'atica
    (CIEM-FaMAF), Conicet. C\'ordoba. Argentina.
    Supported by Secyt-UNC project 33620180100465CB.}
}
\maketitle

\begin{abstract}
  We formalize the basic results on cofinality of linearly ordered
  sets and ordinals and \v{S}anin's Lemma for uncountable families of
  finite sets. We work in the set theory framework of
  Isabelle/ZF, using the Axiom of Choice as needed.
\end{abstract}


\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introduction}

The session we present gathers very basic results built on the set
theory formalization of Isabelle/ZF
\cite{DBLP:journals/jar/PaulsonG96}. In a sense, some of the material
formalized here corresponds to a natural continuation of that
work. This is even clearer after perusing Section~\ref{sec:zf-lib},
where notions like cardinal exponentiation are first defined, together
with various lemmas that do not depend on the Axiom of Choice ($\AC$);
the same holds for the basic theory of cofinality of ordinals, which
is developed in Section~\ref{sec:cofinality}. In
Section~\ref{sec:cardinal-lib}, (un)countability is defined
and several results proved, now using $\AC$ freely; the latter is also
needed to prove König's Theorem on cofinality of cardinal
exponentiation. The simplest infinitary version of the Delta
System Lemma (DSL, also known as the ``Sunflower Lemma'') due to  \v{S}anin
is proved in Section~\ref{sec:dsl}, and it is applied to prove
that Cohen posets satisfy the \emph{countable chain condition}.

A greater part of this development was motivated by an ongoing joint
project on the formalization of the ctm approach to
forcing~\cite{2020arXiv200109715G} by Gunther, Pagano, Steinberg, and
the author. Indeed, most of the results presented here are required
for the development of forcing. As it turns out, the material as
formalized presently will not be part of the forcing formalization,
since for that goal we need relativized versions of both the concepts
and the proofs.

A cross-linked HTML version of the development can be found at
\url{https://cs.famaf.unc.edu.ar/~pedro/Delta_System_Lemma/html}.

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{root}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
